---
title: "PascPhen Report"
output: 
  html_document: 
    highlight: tango
    toc: yes
---


```{r, echo=FALSE, include=FALSE}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(data.table, devtools, backports, Hmisc, tidyr,dplyr,ggplot2,plyr,scales,readr,Rmisc,
               httr, DT, lubridate, tidyverse,reshape2,foreach,doParallel,caret,gbm,lubridate,praznik,
               ggridges, forcats, stats,ppcor, reshape2)

load(paste0(tmp.dir,"/distributions.RData"))


names <- list.files(tmp.dir)
names <- names[names %like% paste0("phen")]
mlho_list<- list()
corr_list <- list()
for (i in 1:length(names)) {
  tryCatch({
    load(paste0(tmp.dir,"/",names[i]))
    mlho_list[[i]] <- features
    dat.i <- data.frame(corrs)
    dat.i$features <- rownames(dat.i)
    rownames(dat.i) <- NULL
    library(reshape2)
    dat.i.wide <- dat.i %>%
  melt(id.var="features") %>%
  arrange(features, variable)
    colnames(dat.i.wide) <- c("var1","var2","correlation")
    dat.i.wide$correlation <- round(dat.i.wide$correlation,3)
    corr_list[[i]] <- dat.i.wide
  }, 
  error = function(fr) {cat("ERROR :",conditionMessage(fr), "\n")})
}

features.table <- do.call(rbind, lapply(mlho_list, data.frame, stringsAsFactors=FALSE))
mlho_list$X <- NULL

write.csv(features.table, 
          file = file.path(output_dir, paste0("table_",siteid,"_",long.thres,"_",substr(long.perc,3,nchar(long.perc)),"_",gsub("\\.", "",data_type),"_", Sys.Date(),".csv")), row.names = FALSE)

cor.tab <- do.call(rbind, lapply(corr_list, data.frame, stringsAsFactors=FALSE))
cor.tab$X <- NULL
```
**This report is generated from  `r siteid` 4Ce data `r data_type` on `r Sys.Date()`.** 

## Exploring the PASC phenotypes

### Distribution of phenotypes by first record
The figure below represents the distribution of the phenotypes over the 6 stages

```{r, echo=FALSE, error=FALSE,fig.align='center', fig.width= 10, warning=FALSE}
pl1
```


### Phenotypes to model
From the initial list, you can pursue phenotyping the following PASC phenotypes.

```{r, echo=FALSE, fig.align='center', fig.width= 10}
datatable(phen.prev, options = list(pageLength = 5), filter = 'bottom')
```
The table presents the `r nrow(phen.prev)` phenotypes for which we have x positives, where x>=`r long.perc` of the total patients with available data from `r long.thres` days or longer after hospitalization/infection

## Applying MLHO to perform PheWAS 
```{r, echo=FALSE, fig.align='center', fig.width= 10, warning=FALSE, error=FALSE, message=FALSE}
pl2
datatable(features.table, options = list(pageLength = 30), filter = 'bottom')
```

correlation matrix among the identified features.
```{r, echo=FALSE, fig.align='center', fig.width= 10}
datatable(cor.tab, options = list(pageLength = 5), filter = 'bottom')
```

## info
Ask questions or report issues: `rebecca.mesa01@universitadipavia.it`

