---
title: "PascPhen Report"
output: 
  html_document: 
    highlight: tango
    toc: yes
---


```{r, echo=FALSE, include=FALSE}
library(pascPhen)
library(mlho)
if(!require(pacman)) install.packages("pacman")
pacman::p_load(data.table, devtools, backports, Hmisc, tidyr,dplyr,ggplot2,plyr,scales,readr,Rmisc,
               httr, DT, lubridate, tidyverse,reshape2,foreach,doParallel,caret,gbm,lubridate,praznik,
               ggridges, forcats, stats, FourCePhase2.1Data, ppcor,pascPhen, reshape2)
```
**This report is generated from  `r siteid` 4Ce data `r data_type` on `r Sys.Date()`.** 

## Exploring the PASC phenotypes

### Distribution of phenotypes by first record
The figure below represents the distribution of the phenotypes over the 6 stages

```{r, echo=FALSE, error=FALSE,fig.align='center', fig.width= 10, warning=FALSE}
load(tmp.file)
pl1
```


### Phenotypes to model
From the initial list, you can pursue phenotyping the following PASC phenotypes.

```{r, echo=FALSE, fig.align='center', fig.width= 10}
datatable(phen.prev, options = list(pageLength = 5), filter = 'bottom')

```
The table presents the `r nrow(phen.prev)` phenotypes for which we have x positives, where x>=`r long.perc` of the total patients with available data from `r long.thres` days or longer after hospitalization/infection

## Applying MLHO to perform PheWAS 
```{r, echo=FALSE, fig.align='center', fig.width= 10, warning=FALSE, error=FALSE, message=FALSE}
pl2
datatable(features.table, options = list(pageLength = 30), filter = 'bottom')
```

correlation matrix among the identified features.
```{r, echo=FALSE, fig.align='center', fig.width= 10}
datatable(cor.tab, options = list(pageLength = 5), filter = 'bottom')

```

## info
Ask questions or report issues: `rebecca.mesa01@universitadipavia.it`

